\papertitle

### 2014-03-17 final-report/paper.tex
graph view changes

The \graph view provides context to the user. It provides a visualization of the physical query plan. A query plan is an ordered set of steps used to access data in database management systems. The physical query plan is represented by a graph where each node represents a query fragment, and each link represents inter-worker communication from one query fragment to the next. Each query fragment is a collection of query operators. One can think of a query fragment as a job unit that gets scheduled on a cluster of machines. Consequently transitioning from one query fragment to another generally incurs all-to-all communication between worker nodes on the cluster.

The user can perform two classes of actions that will render a new visualization in the performance panel:
\begin{itemize}

    \item \textbf{Empty selection:} by default if no fragments or edges are selected in the \graph view, an \overall visualization is rendered in the performance panel. This visualization displays the aggregate worker utilization over time for each fragment and provides cluster utilization data at a glance.
    \item \textbf{Fragment selection:} when selecting a fragment, a \fragment visualization gets displayed in the performance panel. The \fragment visualization provides aggregate cluster utilization information complemented by per-worker task schedules. When a fragment is selected, the operands inside of the fragment in the \graph view are color-coded to allow the user to easily match each task in the per-worker schedule in the performance window with the corresponding query operator.
    \item \textbf{Edge selection:} upon selecting one or more fragment-to-fragment edges, a \network visualization gets displayed, providing aggregate exchange data between cluster nodes.

@@

\textbf{Graph Rendering:} We use D3 to render the \graph graph, which allows us to support various interactions and transistions. We use GraphViz \cite{graphviz} in the backend to generate graph layout information. GraphViz allows us to generate graph layouts optimized for space. The layout information is then fed into a D3-based rendering engine.

\textbf{Graph Navigation:} A query plan can be arbitrarily large. Thus we offer two mechanisms that facilitate exploration of the graph for the user: (1) expanding/reducing fragments and (2) zooming. The first mechanism can reduce the size of the graph by a constant factor by collapsing operators that compose a fragment into a single node. The user can click once on a collapsed fragment to expand it, and click once on an expanded fragment to collapse it. Expanding or reducing a node can cause large changes in the graph layout because GraphViz optimizes the graph layout to minimize area. We implemented transitions to allow the user to track the fragments as they get shuffled around after one fragment gets expanded or reduced. The second mechanism that facilitates exploration of the graph for the user is zooming. This feature was implemented in D3 and allows the user to quickly jump from one part of the graph to another, by zooming in and out of the view.

>>>>

The \graph view allows the user to selectively browse different performance visualizations tied to a given query execution. The \graph view provides as the name indicates, a visualization of the physical query plan. A query plan is an ordered set of steps used to access data in database management systems. The physical query plan is represented by a graph where each node represents a query fragment, and each link represents inter-worker communication from one query fragment to the next. Each query fragment is a collection of query operators. One can think of a query fragment as a job unit that gets scheduled on a cluster of machines. Consequently transitioning from one query fragment to another generally leads to all-to-all communication between worker nodes on the cluster.

The user can perform three classes of actions that will render a new visualization in the performance panel:
\begin{itemize}

    \item \textbf{Empty selection:} by default if no fragments or edges are selected in the \graph view, an \overall visualization is rendered in the performance panel. This visualization displays the aggregate worker utilization over time for each fragment and provides cluster utilization data at a glance.
    \item \textbf{Fragment selection:} when selecting a fragment, a \fragment visualization gets rendered in the performance panel. The \fragment visualization provides aggregate cluster utilization information complemented by per-worker task schedules. When a fragment is selected, the operands inside of the fragment in the \graph view are color-coded to allow the user to easily match each task in the per-worker task schedule in the performance window with the corresponding query operator in the browser window.
    \item \textbf{Edge selection:} upon selecting one or more fragment-to-fragment edges, a \network visualization gets displayed, providing information on inter-worker communication from one query fragment to the next.

@@

\textbf{Graph Rendering:} We use D3 to render the \graph graph, which allows us to support various interactions and transistions. We use GraphViz \cite{graphviz} in the backend to generate graph layout information. GraphViz allows us to generate graph layouts that are optimized to minimize area footprint. The layout information is then fed into a D3-based rendering engine which supports various interactions and animations.

\textbf{Graph Navigation:} A query plan can be arbitrarily large. Thus we offer two mechanisms that facilitate exploration of the graph for the user: (1) expanding/reducing fragments and (2) zooming. The first mechanism can reduce the size of a graph by a constant factor by collapsing operators that compose a fragment into a single fragment node. The user can click once on a collapsed fragment to expand it, and click once on an expanded fragment to collapse it. Expanding or reducing a node can cause large changes in the graph layout as GraphViz changes the graph layout to minimize overall area. To address these layout changes, we implemented transitions to allow the user to track the fragments as those get reshuffled. The second mechanism that facilitates exploration of the graph for the user is zooming. This feature was implemented in D3 and allows the user to quickly jump from one part of the graph to the next, by zooming in and out of the graph view.
